---
title: "Assignments"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

title: "Assignment 1"
author: "Hannah"
date: "9/18/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.

```{r}
#Install data set package

#install.packages("datasets")

#Load the library of datasets 

library(datasets)

```

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

Answer to problem 1: It is useful to rename the dataset because any time you need to analyze the data in any way, you can use the new name to tell the computer to use that data set for the analysis.Also, if you change the data in any way - such as removing outliers - it has to have a new name, so that if someone wants to edit the original data they will not be confused. It is very important for analysis to be replicable, so renaming the data allows you to do what you need to it without taking away someone's else's ability to do their own analysis to the original dataset.

```{r}
#Load USArrests data set

USArrests

#Rename USArrests dat

dat<-USArrests

#Now USArrests data set is called dat
```
### Problem 2

Use this command to make the state names into a new variable called State. 

```{r, eval=FALSE}
dat$state <- tolower(rownames(USArrests))
```

This dataset has the state names as row names, so we just want to make them into a new variable. We also make them all lower case, because that will help us draw a map later - the map function requires the states to be lower case.


List the variables contained in the dataset `USArrests`.

Answer to problem 2: "Murder" "Assault"  "UrbanPop" "Rape" "state"

```{r}
#finding names of state variables

names(dat)
```
### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

Answer: Murder is a quantitative variable. 

What R Type of variable is it?

Answer: In terms of R Types of variables Murder is a numeric variable.


### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

Answer: The information contained in the dataset has to do with 4 categories of crimes and their frequency of arrest rates per 100,000 residents in all 50 states. The numbers refer to how many instances of arrests for that category of crime occurred in that state per every 100,000 residents. 
  

### Problem 5

Draw a histogram of `Murder` with proper labels and title.

```{r}
#Make a histogram of murder with proper axis titles (Problem 5)

hist(dat$Murder, xlab = "Number of Murder Arrests", main = "Histogram of Murder Arrests")
```

### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

Answer to problem 6: The mean of murder is 7.788, the median of murder is 7.250. While mean and median are both measures of central tendency, they have relevant differences. Mean is an average of all of the numbers in the set. It is found by adding up all the numbers in the set and dividing that sum by the total number of variables in the set.The median is the number in the center when you line up all of the variables in numerical order.This means that half of the data lies below the median and the other half above.A quartile represents the three values in a sample that divide the distribution of the data into even fourths.This means that each quartile contains 25% of the data. I think that R gives you the first and third quartiles because these two measures bookend the interquartile range (IQR = Q3-Q1). The interquartile range represents the middle half of the values,and the size of the IQR reveals how spread out the data is. The higher the IQR, the more spread out the data is and vice versa.  


```{r}
#Summarizing Murder quantitatively (Problem 6)

summary(dat$Murder)
```

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r, echo = TRUE, fig.width = 5, fig.height = 8}
#Make a histogram of Assault with the proper axis titles (Problem 7)

hist(dat$Assault , xlab = "Number of Assaults Arrests" , main = "Histogram of Assault Arrests")

#Make a histogram of Rape with proper axis titles (Problem 7)

hist(dat$Rape , xlab = "Number of Rape Arrests" , main = "Histogram of Rape Arrests")

#Make a plot of all 3 histograms (Murder, Assault, and Rape) together (Problem 7)

par(mfrow=c(3,1)) 
hist(dat$Murder, xlab = "Number of Murder Arrests", main = "Histogram of Murder Arrests")
hist(dat$Assault , xlab = "Number of Assault Arrests" , main = "Histogram of Assault Arrests")
hist(dat$Rape , xlab = "Number of Rape Arrests" , main = "Histogram of Rape Arrests")
```
What does the command par do, in your own words (you can look this up by asking R `?par`)?

Answer: The command par is a way in which multiple plots can be viewed together.It essentially set the parameters for viewing graphs. For example, in this problem I wanted to display 3 graphs each with their own variables.The graphs all have the same y-axis (frequency) and all the x-axes are "number of." This command allows the histograms to be viewed simultaneously and thus the data to be easily analyzed and compared.


What can you learn from plotting the histograms together?

Answer: By plotting the histograms together, you can learn about differences and similarities in the arrest rates for these 3 different categories.
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

Run this code:

```{r, eval = FALSE, fig.width = 7.5, fig.height = 4}
library('maps') 
library('ggplot2') 

ggplot(dat, aes(map_id=state, fill=Murder)) + 
  geom_map(map=map_data("state")) + 
  expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```